---
title: "README"
output: html_document
---
First we change the working directory to UCI HAR Dataset 
```{r}
setwd("~/UCI HAR Dataset")
```

Reading the features.txt file
```{r} 
features<-read.table("features.txt")
```

Removing the first column of features which converts features to vector
```{r} 
features<-features[,2]
```

Reading training data
```{r}
setwd("~/UCI HAR Dataset/train")
dt_train<-read.table("X_train.txt", col.names = features)
train_activity<-read.table("y_train.txt")
```
Reading test data
```{r}
setwd("~/UCI HAR Dataset/test")
dt_test<-read.table("X_test.txt",col.names = features)
test_activity<-read.table("y_test.txt")
```
Reading Activity Labels
```{r}
setwd("~/UCI HAR Dataset")
activity_labels<-read.table("activity_labels.txt")
```
Replacing the data in train_activity and test_activity with activity labels so that data becomes more redable 
```{r}
test_activity<-apply(test_activity,1,function(x){
  switch(x,"WALKING","WALKING_UPSTAIRS","WALKING_DOWNSTAIRS","SITTING","STANDING","LAYING")
})

train_activity<-apply(train_activity,1,function(x){
       switch(x,"WALKING","WALKING_UPSTAIRS","WALKING_DOWNSTAIRS","SITTING","STANDING","LAYING")})
```
Creating an activity array with train_activity and test_activity
```{r}
activity<-c(train_activity,test_activity)
```
Merging the train and test data set
```{r}
dt<-rbind(dt_train,dt_test)
```
Adding activity column to our mergerd data set
```{r}
dt$activityname<-activity
```
Extracting only mean and sd columns using grepl function.

We use sapply function over column names of dt and generate a logical vector which have **TRUE** value if "Mean" is present in the column name
```{r}
meanlogi<-sapply(names(dt), function(x)grepl("mean",x,ignore.case = TRUE))
```
Similarly we do the apply same logic to get the logical vector for Standard Deviation
```{r}
stdlogi<-sapply(names(dt), function(x)grepl("std",x,ignore.case = TRUE))
```
Now we generate a logi vector which have **TRUE** value for all mean and sd columns

```{r}
logi<-as.logical()
for (i in 1:562) {
  logi[i]<-meanlogi[i]||stdlogi[i]
}

```
Setting the last column to TRUE (because it has our activity information) and generating the tidy data 
```{r}
 logi[562]<-TRUE
 tidydata<-dt[logi]
```
To calculate the Average according to activity we need to group the data. So using split function to group the data
```{r}
sp<-split(tidydata,factor(tidydata[,87]))
```
Calculating the average 
```{r}
 avgact<-data.frame()

 for (i in seq_along(sp)) {
  avgact[i,1]<-names(sp)[[i]]
   for (j in 1:86) {
     
     avgact[i,j+1]<-colMeans(sp[[i]][j])
   }
   
 }
colnames(avgact)<-c('Activity',names(tidydata)[1:86])
```
Saving the avgact to final.txt file
```{r}
write.table(avgact,"final.txt",row.names = FALSE)
```

